### 2024-09-18 manuscript/manuscript.tex
Combine table 7 and 8

%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}

@@

\textbf{Authors}:\\
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Pierre Legagneux\textsuperscript{3,4}, Gilles Gauthier\textsuperscript{5}, Éliane Duchesne\textsuperscript{1}, Marie-Christine Cadieux \textsuperscript{3}, Dominique Berteaux\textsuperscript{1}, Dominique Fauteux\textsuperscript{6} and Joël Bêty\textsuperscript{1}\\[1.5em]

\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France \\
\textsuperscript{5} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{6} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada
\\[1.5em]

\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]

@@

Modeling ecological communities requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community. However, accurately determining the abundance of all species is rarely achievable. Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories. Additionally, food webs consist of species with varying body sizes depending on their trophic level, with top-level species often being highly mobile and having large home ranges \citep{mccann2005}. Therefore, community models must use landscape-wide estimates of species abundance to accurately represent trophic fluxes. Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.

The composition of ecological communities is influenced by various factors acting at different temporal and spatial scales, leading to noisy data and emphasizing the need for long-term data sets \citep{magurran2010, lindenmayer2012}. Species abundances are influenced by stochastic effects \citep{hubbell2001}, environmental changes (e.g., climate warming), and species interactions, contributing to data variability. For instance, the composition of a community could be driven simultaneously by the intra-annual seasonal variations in the environment, the multi-year cyclic dynamics of an important prey or predator and a relatively slow but directional climate warming. Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.

@@

\subsubsection*{Rough-legged hawk, peregrine falcon and common raven}
Peregrine falcons, rough-legged hawks and common ravens nest on cliffs, near ravines, and on large rocky outcrops and tend to reuse the same nesting sites from one year to the next \citep{beardsell2016}. Systematic monitoring of every known or potential nesting site has been carried out in the Qarlikturvik valley, Black plateau and Southern plateau since 2007 and throughout the study area since 2013 \citep{beardsell2016, gauthier2020avianpred}. Observers walked along ridges and scan surrounding areas from vantage points to detect nesting birds. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. They can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species. Each year the observers use slightly different paths to sample the areas, but locate the nests in the same positions, which supports a high probability of detection for these species. Most nesting sites were located in the upland zones of the study area, which include the Black Plateau, Southern Plateau and Camp 3 Plateau.

@@

\input{tables/table_summary_methods.tex}
%label-> table:summary_methods
\newpage

@@

\section{Species abundance}

\subsection{Uncertainty on estimates}
We did not quantitatively assessed uncertainty on each estimate of species abundance. However, we provide a qualitative measure of uncertainty based on the level of confidence derived from field experts, and in some cases, the fit of statistical models to estimate density (\textbf{Table \ref{table:confidence_estimate_abundance}}).

\input{tables/confidence_estimate_abundance.tex}
%label -> table:confidence_estimate_abundance
\newpage

\subsection{Abundance estimates}
\input{tables/species_mean_abudance.tex}
%label -> table:species_mean_abudance
\newpage

The following figures present the time series of species abundances. If different methods were used, we present the estimates derived from each method with different colors (see \textbf{Table \ref{table:confidence_estimate_abundance}} for the corresponding method and confidence in estimates).

>>>>

%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
\usepackage{pdflscape}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}
\usepackage{geometry}

@@

\textbf{Authors}:\\
Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1},
Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]

\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]

\textbf{Corresponding Authors}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]

@@

Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community. However, accurately determining the abundance of all species is rarely achievable. Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories. Additionally, food webs consist of species with varying body sizes depending on their trophic level, with top-level species often being highly mobile and having large home ranges \citep{mccann2005}. Therefore, community models must use landscape-wide estimates of species abundance to accurately represent trophic fluxes. Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.

The composition of ecological communities is influenced by various factors acting at different temporal and spatial scales, leading to noisy data and emphasizing the need for long-term data sets \citep{magurran2010, lindenmayer2012}. Species abundances are influenced by stochastic effects \citep{hubbell2001}, environmental changes (e.g., climate warming), and species interactions, contributing to data variability. For instance, the composition of a community could be driven simultaneously by intra-annual seasonal variations, multi-year cyclic variations (e.g., El Niño) and slow but directional long-term variations in the environment \citep{brown1990, snyder2006}. Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.

@@

\subsubsection*{Rough-legged hawk, peregrine falcon and common raven}
Peregrine falcons, rough-legged hawks and common ravens nest on cliffs, near ravines, and on large rocky outcrops and tend to reuse the same nesting sites from one year to the next \citep{beardsell2016}. Systematic monitoring of every known or potential nesting site has been carried out in the Qarlikturvik valley, Black plateau and Southern plateau since 2007 and throughout the study area since 2013 \citep{beardsell2016, gauthier2020avianpred}. Observers walked along ridges and scanned surrounding areas from vantage points to detect nesting birds. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. They can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species. Each year the observers use slightly different paths to sample the areas, but locate the nests in the same positions, which supports a high probability of detection for these species. Most nesting sites were located in the upland zones of the study area, which include the Black Plateau, Southern Plateau and Camp 3 Plateau.

@@

@@

\newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
\begin{landscape}
\section{Species abundance}
\input{tables/table_summary_methods.tex}
%label-> table:summary_methods
\end{landscape}
\restoregeometry
\newpage

\input{tables/species_mean_abudance.tex}
%label -> table:species_mean_abudance
\newpage

### 2024-09-18 manuscript/tables/species_mean_abudance.tex
uniformize species name in the final data set
Combine table 7 and 8

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:29 2024
\begin{table}[ht]
\centering
\caption{Mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2). In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty.} 
\label{table:species_mean_abudance}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllrrr}
  \hline
Species & Method & Mean ± sd & n & Minimum & Maximum \\ 
  \hline
king eider & nest sampling (extrapolation habitat) & 25 &  &  &  \\ 
  king eider & incidental observations (relative abundance) & 106 &  &  &  \\ 
  long-tailed duck & nest sampling (extrapolation habitat) & 20 &  &  &  \\ 
  long-tailed duck & incidental observations (relative abundance) & 191 &  &  &  \\ 
  Pacific loon & nest sampling & 4 ± 3 &   4 &   0 &   6 \\ 
  red-throated loon & nest sampling & 64 ± 15 &   4 &  42 &  76 \\ 
  cackling goose & nest sampling & 50 ± 59 &  28 &   2 & 214 \\ 
  snow goose & combined methods & 25289 ± 13327 &  23 & 2505 & 49076 \\ 
  tundra swan & nest sampling & 1 ± 1 &   4 &   0 &   2 \\ 
  rough-legged hawk & nest sampling & 25 ± 24 &  16 &   0 &  66 \\ 
  peregrine falcon & nest sampling & 10 ± 1 &   8 &   8 &  12 \\ 
  snowy owl & nest sampling & 16 ± 33 &  26 &   0 & 144 \\ 
  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  &  \\ 
  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  &  \\ 
  American golden-plover & distance sampling & 774 ± 275 &   8 & 306 & 1040 \\ 
  black-bellied plover & transects (relative abundance) & 22 &  &  &  \\ 
  black-bellied plover & incidental observations (relative abundance) & 61 &  &  &  \\ 
  common-ringed plover & nest sampling & 55 ± 9 &   3 &  44 &  62 \\ 
  ruddy turnstone & transects (relative abundance) & 40 &  &  &  \\ 
  ruddy turnstone & incidental observations (relative abundance) & 54 &  &  &  \\ 
  red knot & transects (relative abundance) & 64 &  &  &  \\ 
  red knot & incidental observations (relative abundance) & 237 &  &  &  \\ 
  pectoral sandpiper & transects (relative abundance) & 79 &  &  &  \\ 
  pectoral sandpiper & incidental observations (relative abundance) & 260 &  &  &  \\ 
  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  &  \\ 
  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  &  \\ 
  white-rumped sandpiper & incidental observations (relative abundance) & 1156 &  &  &  \\ 
  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  &  \\ 
  buff-breasted sandpiper & incidental observations (relative abundance) & 8 &  &  &  \\ 
  red phalarope & transects (relative abundance) & 156 &  &  &  \\ 
  red phalarope & incidental observations (relative abundance) & 276 &  &  &  \\ 
  glaucous gull & nest sampling & 73 ± 7 &  17 &  59 &  80 \\ 
  long-tailed jaeger & nest sampling & 272 ± 285 &  17 &   0 & 900 \\ 
  parasitic jaeger & banding & 17 &  &  &  \\ 
  parasitic jaeger & nest sampling & 8 &  &  &  \\ 
  common raven & transects (relative abundance) & 14 &  &  &  \\ 
  common raven & incidental observations (relative abundance) & 31 &  &  &  \\ 
  horned lark & transects (relative abundance) & 362 &  &  &  \\ 
  horned lark & incidental observations (relative abundance) & 411 &  &  &  \\ 
  American pipit & transects (relative abundance) & 53 &  &  &  \\ 
  American pipit & incidental observations (relative abundance) & 87 &  &  &  \\ 
  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  &  \\ 
  snow bunting & transects (relative abundance) & 18 &  &  &  \\ 
  snow bunting & incidental observations (relative abundance) & 276 &  &  &  \\ 
  brown lemming & trapping & 54043 ± 93530 &  27 &   0 & 447630 \\ 
  collared lemming & trapping & 8128 ± 10334 &  27 &   0 & 39302 \\ 
  Arctic hare & incidental observations (relative abundance) & 6 &  &  &  \\ 
  ermine & testimonials & 40 ± 37 &  27 &   8 & 156 \\ 
  Arctic fox & individual home range & 53 &  &  &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

>>>>

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Sep 10 14:57:19 2024
\begin{table}[ht]
\centering
\caption{Selected methods to represent the mean (± standard deviation) of annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).} 
\label{table:species_mean_abudance}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{0.85\textwidth}{llllr}
  \hline
Species & Method & Mean & sd & n \\ 
  \hline
king eider & incidental observations (relative abundance) & 106 &  &  \\ 
  long-tailed duck & incidental observations (relative abundance) & 191 &  &  \\ 
  Pacific loon & nest sampling & 4 & 3 &   4 \\ 
  red-throated loon & nest sampling & 64 & 15 &   4 \\ 
  cackling goose & nest sampling & 140 & 41 &   7 \\ 
  snow goose & combined methods & 31852 & 12092 &  12 \\ 
  tundra swan & nest sampling & 1 & 1 &   4 \\ 
  rough-legged hawk & nest sampling & 25 & 24 &  16 \\ 
  peregrine falcon & nest sampling & 10 & 1 &   8 \\ 
  snowy owl & nest sampling & 16 & 33 &  26 \\ 
  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  \\ 
  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  \\ 
  American golden-plover & distance sampling & 774 & 275 &   8 \\ 
  black-bellied plover & transects (relative abundance) & 22 &  &  \\ 
  common-ringed plover & nest sampling & 55 & 9 &   3 \\ 
  ruddy turnstone & transects (relative abundance) & 40 &  &  \\ 
  red knot & transects (relative abundance) & 64 &  &  \\ 
  pectoral sandpiper & transects (relative abundance) & 79 &  &  \\ 
  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  \\ 
  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  \\ 
  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  \\ 
  red phalarope & transects (relative abundance) & 156 &  &  \\ 
  glaucous gull & nest sampling & 73 & 7 &  17 \\ 
  long-tailed jaeger & nest sampling (extrapolation habitat) & 272 & 285 &  17 \\ 
  parasitic jaeger & banding & 17 &  &  \\ 
  common raven & incidental observations (relative abundance) & 31 &  &  \\ 
  horned lark & incidental observations (relative abundance) & 411 &  &  \\ 
  American pipit & incidental observations (relative abundance) & 87 &  &  \\ 
  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  \\ 
  snow bunting & incidental observations (relative abundance) & 276 &  &  \\ 
  brown lemming & trapping & 54043 & 93530 &  27 \\ 
  collared lemming & trapping & 8128 & 10334 &  27 \\ 
  Arctic hare & incidental observations (relative abundance) & 6 &  &  \\ 
  ermine & testimonials & 40 & 37 &  27 \\ 
  Arctic fox & individual home range & 53 &  &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}
